---
title: "Task1"
author: "Pavitra Desai"
date: "12/08/2021"
output: 
  html_document:
    keep_md: yes
---

# GRIP under The Sparks Foundation - Data Science and Business Analytics
## Prediction using Supervised ML  
### **Problem Statement** - Predict the percentage Scores of an student based on the no. of study Hours using Simple Linear Regression
  
**Importing necessary Libraries**
```{r}
# Uncomment the following to install packages 
#install.packages("ggplot2")
#install.packages("caTools")
library(ggplot2)
library(caTools)
```
  
**Downloading, understanding and cleaning dataset**  
Data can be downloaded from http://bit.ly/w-data
```{r}
url<-"http://bit.ly/w-data"
data<-read.csv(file=url)
head(data)
dim(data)
summary(data)
is.null(data)
any(is.na(data))
```
  Data consists of 25 rows, 2 columns ; Hours of study per day(independent variable) and Score(dependent variable) . 
  Data is clean with no null or missing values.  
  
**Data Visualization**
```{r}
plot(data)
```

Subsetting data into training sets(75%) and testing sets(25%) with split ratio=4/5
```{r}
set.seed(100)
split<-sample.split(data$Hours,SplitRatio = 4/5)
train<-subset(x=data,split==TRUE)
test<-subset(x=data,split==FALSE)
model<-lm(Scores~Hours,train)
summary(model)
```
  Plotting Training set 
```{r}
ggplot() +
  geom_point(aes(x=train$Hours,y=train$Scores),colour="red")+
  geom_line(aes(x=train$Hours,y=predict(model,train)),colour="green")+
  ggtitle ("Training set")+
  xlab("Hours")+
  ylab("Scores")
```
  
Testing Set Results 
```{r}
ggplot() +
  geom_point(aes(x=test$Hours, y=test$Scores),colour = 'blue') +
  geom_line (aes( x= train$Hours, y=predict(model,train)),colour = 'green') +
  ggtitle ('Testing Set') +
  xlab ('Hours') +
  ylab ('Scores') 
```
   
  Comparing Actual and Predicted values
```{r}
predictions<-predict(model,test)
new<-cbind(test$Scores,as.integer(predictions))
colnames(new)<-c("Actual","Predicted")
new<-as.data.frame(new)
new
```

Simple Linear Regression Model to predict unknown values
```{r}
linear_model<-lm(formula = Scores~Hours,data)
plot(data)
abline(linear_model,col="red")
test<-data.frame(Hours=9.25)
result<-predict(linear_model,test)
points(test,result,col="blue",pch=8)
```
```{r , echo=FALSE}
sprintf("If a student studies for 9.25 hours/day he will Score %f",result)
```
  
**Root Mean Squared Error (RMSE) to measure the model prediction error**
```{r}
mse <- mean((new$Actual -new$Predicted)^2)
rmse <- sqrt(mse)
rmse
```
The lower the RMSE, the better the model.